---
title: "Untitled"
author: "João Victor"
date: "2022-10-22"
output: html_document
---

Although the regression was made on my own, most of the pre processing data and EDA was based on this work: <https://www.kaggle.com/code/erikbruin/house-prices-lasso-xgboost-and-a-detailed-eda>  

```{r}
library(tidyverse)
library(corrplot)
library(plyr)
library(gridExtra)
library(caret)
library(mltools)
library(ggrepel)
library(car)
library(glmnet)
library(caTools)
library(e1071)
library(knitr)
library(Metrics)
```

### Read the Data Sets
```{r}
train = read.csv('train.csv')
test = read.csv('test.csv')
```
##### Remove Id Variable
```{r}
test$Id = NULL
train$Id = NULL
```
##### Merge the train and test data set
```{r}
test$SalePrice = NA
train_x = rbind(train, test)
```

##### SalePrice Histogram
```{r}
ggplot(train_x[!is.na(train_x$SalePrice), ], aes(SalePrice)) +
  geom_histogram(aes(y = ..density..), bins = 150, colour = 'cadetblue1', fill = 'skyblue') + 
  geom_density(lwd = 1.2, linetype = 1, colour = 2) + 
  ggtitle('SalePrice Histogram') + theme(plot.title = element_text(hjust = 0.5))
```

##### Correlation Matrix
```{r, }
numericVars = which(sapply(train_x, is.numeric))
numericVarNames = names(numericVars)

train_x_numVars = train_x[, numericVars]
cor_numVars = cor(train_x_numVars, use = 'pairwise.complete.obs') 

cor_sorted = as.matrix(sort(cor_numVars[, 'SalePrice'], decreasing = T))

CorHigh = names(which(apply(cor_sorted, 1, function(x) abs(x) > 0.4)))
cor_numVars = cor_numVars[CorHigh, CorHigh]
corrplot.mixed(cor_numVars, tl.col = 'black', tl.pos = 'lt', )
```

This first correlation matrix were made to see the correlation of the numeric variables in the data set. GarageCars and GarageArea has the highest correlation between the variables, the third highest is TotRmsAbvGrd with GrLivArea, which makes sense for both of them. The number of cars that the garages can support is influenced by the area of the garages. Similarly, the numbers of rooms above grade is influenced by the groud live area.  
We can also see other meaningful correlations: GarageYrBlt with YearBuilt (0.82, the secons highest by the way); and X1stFlrSF with TotalBsmtSf.  The highest correlations with the dependent variable - SalePrice - is OverallQual and GrLivArea.

##### SalePrice Boxplot
```{r}
ggplot(train_x[!is.na(train_x$SalePrice), ], aes(x = factor(OverallQual), y = SalePrice)) + 
  geom_boxplot(col = 'red') + labs(x = 'Overall Quality') + 
  scale_y_continuous(breaks = seq(0, 800000, by = 100000)) + 
  ggtitle('SalePrice Boxplot based on Quality') + theme(plot.title = element_text(hjust = 0.5))
```

##### SalePrice vs GrLivArea Scatter Plot
```{r}
ggplot(train_x, aes(x = GrLivArea, y = SalePrice)) +
  geom_point(col = 'blue') + geom_smooth(method = 'lm', se = F, color = 'black', aes(group = 1)) + 
  scale_y_continuous(breaks = seq(0, 800000, by = 100000)) + 
  ggtitle('SalePrice vs GrLivArea Scatter Plot') + theme(plot.title = element_text(hjust = 0.5))
```

##### Checking for missing values
```{r}
NAcol = colnames(train_x)[colSums(is.na(train_x)) > 0]
sort(colSums(sapply(train_x[NAcol], is.na)), decreasing = T)
```

##### Fixing it
```{r}
#PoolQC
train_x$PoolQC[is.na(train_x$PoolQC)] = 'No Pool'

#MiscFeature
train_x$MiscFeature[is.na(train_x$MiscFeature)] = 'None'

#Alley
train_x$Alley[is.na(train_x$Alley)] = 'No Alley'

#Fence
train_x$Fence[is.na(train_x$Fence)] = 'No Fence'

#FireplaceQu
train_x$FireplaceQu[is.na(train_x$FireplaceQu)] = 'No Fireplace'

#LotFrontage
train_x$LotFrontage[is.na(train_x$LotFrontage)] = median(train_x$LotFrontage, na.rm = T)

#GarageType, GarageFinish, GarageQual, GarageCond
train_x = train_x %>%
  mutate_at(c('GarageType', 'GarageFinish', 'GarageQual', 'GarageCond'), ~replace_na(., 'No Garage'))

#GarageYrBlt
train_x$GarageYrBlt[is.na(train_x$GarageYrBlt)] = train_x$YearBuilt[is.na(train_x$GarageYrBlt)]
train_x[train_x$GarageYrBlt == 2207, ]$GarageYrBlt = 2007

#BsmtQual, BsmtCond, BsmtExposure, BsmtFinType1, BsmtFinType2
train_x = train_x %>%
  mutate_at(c('BsmtQual', 'BsmtCond', 'BsmtExposure', 'BsmtFinType1', 'BsmtFinType2'), ~replace_na(., 'No Basement'))

#MasVnrType
train_x$MasVnrType[is.na(train_x$MasVnrType)] = 'None'

#MasVnrArea
train_x$MasVnrArea[is.na(train_x$MasVnrArea)] = median(train_x$MasVnrArea, na.rm = T)

#Electrical
train_x$Electrical[is.na(train_x$Electrical)] = 'SBrkr'

#MSZoning
train_x$MSZoning[is.na(train_x$MSZoning)] = median(train_x$MSZoning, na.rm = T)

#BsmtFullBath
train_x$BsmtFullBath[is.na(train_x$BsmtFullBath)] = median(train_x$BsmtFullBath, na.rm = T)

#BsmtHalfBath
train_x$BsmtHalfBath[is.na(train_x$BsmtHalfBath)] = median(train_x$BsmtHalfBath, na.rm = T)

#Functional
train_x$Functional[is.na(train_x$Functional)] = median(train_x$Functional, na.rm = T)

#Exterior1st
train_x$Exterior1st[is.na(train_x$Exterior1st)] = names(sort(-table(train_x$Exterior1st))) [1]

#Exterior2nd
train_x$Exterior2nd[is.na(train_x$Exterior2nd)] = names(sort(-table(train_x$Exterior2nd))) [1]

#BsmtFinSF1
train_x$BsmtFinSF1[is.na(train_x$BsmtFinSF1)] = 0

#BsmtFinSF2
train_x$BsmtFinSF2[is.na(train_x$BsmtFinSF2)] = 0

#BsmtUnfSF
train_x$BsmtUnfSF[is.na(train_x$BsmtUnfSF)] = 0

#TotalBsmtSF
train_x$TotalBsmtSF[is.na(train_x$TotalBsmtSF)] = 0

#KitchenQual
train_x$KitchenQual[is.na(train_x$KitchenQual)] = 'TA' #replace with most common value

#GarageCars
train_x$GarageCars[is.na(train_x$GarageCars)] = 0

#GarageArea
train_x$GarageArea[is.na(train_x$GarageArea)] = 0

#SaleType
train_x$SaleType[is.na(train_x$SaleType)] = names(sort(-table(train_x$SaleType)))[1]
```

##### Check NA values again
```{r}
colnames(train_x)[colSums(is.na(train_x)) > 0]
```

```{r}
## Drop Utilities
train_x$Utilities = NULL
```

##### Create House Variables
```{r}
train_x$Remod = as.factor(ifelse(train_x$YearBuilt == train_x$YearRemodAdd, 0, 1)) #0 = No remodeling, 1 = Remodeling
train_x$IsNew <- as.factor(ifelse(train_x$YrSold == train_x$YearBuilt, 1, 0)) #0 = No, 1 = Yes
```

##### Converting Character Variables to Factor
```{r}
str(train_x)

## MSSubClass
train_x$MSSubClass = factor(train_x$MSSubClass, levels = c('20', '30', '40', '45', '50', '60', '70', '75', '80', '85', '90',
                                                           '120', '150', '160', '180', '190'),
                            labels = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15))

##MSZoning
train_x$MSZoning = factor(train_x$MSZoning, levels = c('C (all)', 'RM', 'RL', 'RH', 'FV'),
                          labels = c(0, 1, 2, 3, 4))

##Street
train_x$Street = factor(train_x$Street, levels = c('Pave', 'Grvl'), labels = c(0, 1))

##Alley
train_x$Alley = factor(train_x$Alley, levels = c('No Alley', 'Pave', 'Grvl'), labels =c(0, 1, 2))

##LotShape
train_x$LotShape = factor(train_x$LotShape, levels = c('IR3', 'IR2', 'IR1', 'Reg'), labels = c(0, 1, 2, 3))

##LandContour
train_x$LandContour = factor(train_x$LandContour, levels = c('Low', 'HLS', 'Bnk', 'Lvl'),
                             labels = c(0, 1, 2, 3))

##LotConfig
train_x$LotConfig = factor(train_x$LotConfig, levels = c('FR3', 'FR2', 'CulDSac', 'Corner', 'Inside'),
                           labels = c(0, 1, 2, 3, 4))

##LandSlope
train_x$LandSlope = factor(train_x$LandSlope, levels = c('Sev', 'Mod', 'Gtl'),
                           labels = c(0, 1, 2))

##Neighborhood
#Check where is the better Neighborhood
nb1 <- ggplot(train_x, aes(x=reorder(Neighborhood, SalePrice, FUN=median), y=SalePrice)) +
  geom_bar(stat='summary', fun = "median", fill='blue') + labs(x='Neighborhood', y='Median SalePrice') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks= seq(0, 800000, by=50000)) +
  geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=3) +
  geom_hline(yintercept=163000, linetype="dashed", color = "red") #dashed line is median SalePrice

nb2 <- ggplot(train_x, aes(x=reorder(Neighborhood, SalePrice, FUN=mean), y=SalePrice)) +
  geom_bar(stat='summary', fun = "mean", fill='blue') + labs(x='Neighborhood', y="Mean SalePrice") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks= seq(0, 800000, by=50000)) +
  geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=3) +
  geom_hline(yintercept=163000, linetype="dashed", color = "red") #dashed line is median SalePrice

grid.arrange(nb1, nb2)

##Create NeighRich
train_x$NeighRich[train_x$Neighborhood %in% c('StoneBr', 'NridgHt', 'NoRidge')] = 2
train_x$NeighRich[!train_x$Neighborhood %in% c('MeadowV', 'IDOTRR', 'BrDale', 'StoneBr', 'NridgHt', 'NoRidge')] = 1
train_x$NeighRich[train_x$Neighborhood %in% c('MeadowV', 'IDOTRR', 'BrDale')] = 0
train_x$NeighRich = as.factor(train_x$NeighRich)

#train_x$Neighborhood = factor(train_x$Neighborhood, levels = c('Veenker', 'Timber', 'StoneBr', 'Somerst', 'SawyerW', 'Sawyer',
#                                                               'SWISU', 'OldTown', 'NWAmes', 'NridgHt', 'NPkVill', 'NoRidge',
#                                                               'NAmes', 'Mitchel', 'MeadowV', 'IDOTRR', 'Gilbert', 'Edwards',
#                                                               'Crawfor', 'CollgCr', 'ClearCr', 'BrkSide', 'BrDale', 'Blueste', 'Blmngtn' ),
#                              labels = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24))

##Remove Neighborhood
train_x$Neighborhood = NULL

##Condition1
train_x$Condition1 = factor(train_x$Condition1, levels = c('RRAe', 'RRNe', 'PosA', 'PosN', 'RRAn', 'RRNn', 'Norm', 'Feedr', 'Artery'),
                            labels = c(0, 1, 2, 3, 4, 5, 6, 7, 8))

##Condition2
train_x$Condition2 = factor(train_x$Condition2, levels = c('RRAe', 'RRNe', 'PosA', 'PosN', 'RRAn', 'RRNn', 'Norm', 'Feedr', 'Artery'),
                            labels = c(0, 1, 2, 3, 4, 5, 6, 7, 8))

##BldgType
train_x$BldgType = factor(train_x$BldgType, levels = c('Twnhs', 'TwnhsE', 'Duplex', '2fmCon', '1Fam'),
                          labels = c(0, 1, 2, 3, 4))

##HouseStyle
train_x$HouseStyle = factor(train_x$HouseStyle, levels = c('SLvl', 'SFoyer', '2.5Unf', '2.5Fin', '2Story', '1.5Unf', '1.5Fin', '1Story'),
                            labels = c(0, 1, 2, 3, 4, 5, 6, 7))

##RoofStyle
train_x$RoofStyle = factor(train_x$RoofStyle, levels = c('Shed', 'Mansard', 'Hip', 'Gambrel', 'Gable', 'Flat'),
                           labels = c(0, 1, 2, 3, 4, 5))

##RoofMatl
train_x$RoofMatl = factor(train_x$RoofMatl, levels = c('WdShngl', 'WdShake', 'Tar&Grv', 'Roll', 'Metal', 'Membran', 'CompShg', 'ClyTile'),
                          labels = c(0, 1, 2, 3, 4, 5, 6, 7))

##Exterior1st
train_x$Exterior1st = factor(train_x$Exterior1st, levels = c('WdShing', 'Wd Sdng', 'VinylSd', 'Stucco', 'Stone', 'PreCast',
                                                             'Plywood', 'Other', 'MetalSd', 'ImStucc', 'HdBoard', 'CemntBd',
                                                             'CBlock', 'BrkFace', 'BrkComm', 'AsphShn', 'AsbShng'),
                             labels = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16))	
	
##Exterior2nd
train_x$Exterior2nd = factor(train_x$Exterior2nd, levels = c('Wd Shng', 'Wd Sdng', 'VinylSd', 'Stucco', 'Stone', 'PreCast',
                                                             'Plywood', 'Other', 'MetalSd', 'ImStucc', 'HdBoard', 'CmentBd',
                                                             'CBlock', 'BrkFace', 'Brk Cmn', 'AsphShn', 'AsbShng'),
                             labels = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16))		

##MasVnrType
train_x$MasVnrType = factor(train_x$MasVnrType, levels = c('Stone', 'None', 'CBlock', 'BrkFace', 'BrkCmn'),
                            labels = c(0, 1, 2, 3, 4))

##ExterQual
train_x$ExterQual = factor(train_x$ExterQual, levels = c('Po', 'Fa', 'TA', 'Gd', 'Ex'),
                           labels = c(0, 1, 2, 3, 4))

##ExterCond
train_x$ExterCond = factor(train_x$ExterCond, levels = c('Po', 'Fa', 'TA', 'Gd', 'Ex'),
                           labels = c(0, 1, 2, 3, 4))

##Foundation
train_x$Foundation = factor(train_x$Foundation, levels = c('Wood', 'Stone', 'Slab', 'PConc', 'CBlock', 'BrkTil'),
                            labels = c(0, 1, 2, 3, 4, 5))

##BsmtQual
train_x$BsmtQual = factor(train_x$BsmtQual, levels = c('No Basement', 'Po', 'Fa', 'TA', 'Gd', 'Ex'),
                          labels = c(0, 1, 2, 3, 4, 5))

##BsmtCond
train_x$BsmtCond = factor(train_x$BsmtCond, levels = c('No Basement', 'Po','Fa', 'TA', 'Gd', 'Ex'),
                          labels = c(0, 1, 2, 3, 4, 5))

##BsmtExposure
train_x$BsmtExposure = factor(train_x$BsmtExposure, levels = c('No Basement', 'No', 'Mn', 'Av', 'Gd'),
                              labels = c(0, 1, 2, 3, 4))

##BsmtFinType1
train_x$BsmtFinType1 = factor(train_x$BsmtFinType1, levels = c('No Basement', 'Unf', 'LwQ', 'Rec', 'BLQ', 'ALQ', 'GLQ'),
                              labels = c(0, 1, 2, 3, 4, 5, 6))

##BsmtFinType2
train_x$BsmtFinType2 = factor(train_x$BsmtFinType2, levels = c('No Basement', 'Unf', 'LwQ', 'Rec', 'BLQ', 'ALQ', 'GLQ'),
                              labels = c(0, 1, 2, 3, 4, 5, 6))

##Heating
train_x$Heating = factor(train_x$Heating, levels = c('Wall', 'OthW', 'Grav', 'GasW', 'GasA', 'Floor'),
                         labels = c(0, 1, 2, 3, 4, 5))

##HeatingQC
train_x$HeatingQC = factor(train_x$HeatingQC, levels = c('Po', 'Fa', 'TA', 'Gd', 'Ex'),
                           labels = c(0, 1, 2, 3, 4))

##CentralAir
train_x$CentralAir = factor(train_x$CentralAir, levels = c('N', 'Y'),
                            labels = c(0, 1))

##Electrical
train_x$Electrical = factor(train_x$Electrical, levels = c('Mix', 'FuseP', 'FuseF', 'FuseA', 'SBrkr'),
                            labels = c(0, 1, 2, 3, 4))

##KitchenQual
train_x$KitchenQual = factor(train_x$KitchenQual, levels = c('Po', 'Fa', 'TA', 'Gd', 'Ex'),
                             labels = c(0, 1, 2, 3, 4))

##Functional
train_x$Functional = factor(train_x$Functional, levels = c('Sal', 'Sev', 'Maj2', 'Maj1', 'Mod', 'Min2', 'Min1', 'Typ'),
                            labels = c(0, 1, 2, 3, 4, 5, 6, 7))

##FireplaceQu
train_x$FireplaceQu = factor(train_x$FireplaceQu, levels = c('No Fireplace', 'Po', 'Fa', 'TA', 'Gd', 'Ex'),
                             labels = c(0, 1, 2, 3, 4, 5))

##GarageType
train_x$GarageType = factor(train_x$GarageType, levels = c('No Garage', 'Detchd', 'CarPort', 'BuiltIn', 'Basment', 'Attchd', '2Types'),
                            labels = c(0, 1, 2, 3, 4, 5, 6))

##GarageFinish
train_x$GarageFinish = factor(train_x$GarageFinish, levels = c('No Garage', 'Unf', 'RFn', 'Fin'),
                             labels = c(0, 1, 2, 3))

##GarageQual
train_x$GarageQual = factor(train_x$GarageQual, levels = c('No Garage', 'Po', 'Fa', 'TA', 'Gd', 'Ex'),
                            labels = c(0, 1, 2, 3, 4, 5))

##GarageCond
train_x$GarageCond = factor(train_x$GarageCond, levels = c('No Garage', 'Po', 'Fa', 'TA', 'Gd', 'Ex'),
                            labels = c(0, 1, 2, 3, 4, 5))

##PavedDrive
train_x$PavedDrive = factor(train_x$PavedDrive, levels = c('N', 'P', 'Y'),
                            labels = c(0, 1, 2))

##PoolQC
train_x$PoolQC = factor(train_x$PoolQC, levels = c('No Pool', 'Fa', 'TA', 'Gd', 'Ex'),
                        labels = c(0, 1, 2, 3, 4))

##Fence
train_x$Fence = factor(train_x$Fence, levels = c('No Fence', 'MnWw', 'GdWo', 'MnPrv', 'GdPrv'),
                       labels = c(0, 1, 2, 3, 4))

##MiscFeature
train_x$MiscFeature = factor(train_x$MiscFeature, levels = c('None', 'TenC', 'Shed', 'Othr', 'Gar2', 'Elev'),
                             labels = c(0, 1, 2, 3, 4, 5))

##SaleType
train_x$SaleType = factor(train_x$SaleType, levels = c('Oth', 'ConLD', 'ConLI', 'ConLw', 'Con', 'COD', 'New', 'VWD', 'CWD', 'WD'),
                          labels = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9))

##SaleCondition
train_x$SaleCondition = factor(train_x$SaleCondition, levels = c('Partial', 'Family', 'Alloca', 'AdjLand', 'Abnorml', 'Normal'),
                               labels = c(0, 1, 2, 3, 4, 5))

```

##### Check Again
```{r}
str(train_x)
```

```{r}
colnames(train_x)[colSums(is.na(train_x)) > 0]
```

##### Converting some Numeric Variables to Factor
```{r}
##OverallQual
train_x$OverallQual = as.factor(train_x$OverallQual)

##OverallCond
train_x$OverallCond = as.factor(train_x$OverallCond)

##BsmtFullBath
train_x$BsmtFullBath = as.factor(train_x$BsmtFullBath)

##BsmtHalfBath
train_x$BsmtHalfBath = as.factor(train_x$BsmtHalfBath)

##FullBath
train_x$FullBath = as.factor(train_x$FullBath)

##HalfBath
train_x$HalfBath = as.factor(train_x$HalfBath)

##BedroomAbvGr
train_x$BedroomAbvGr = as.factor(train_x$BedroomAbvGr)

##KitchenAbvGr
train_x$KitchenAbvGr = as.factor(train_x$KitchenAbvGr)

##TotRmsAbvGrd
train_x$TotRmsAbvGrd = as.factor(train_x$TotRmsAbvGrd)

##Fireplaces
train_x$Fireplaces = as.factor(train_x$Fireplaces)

##GarageCars
train_x$GarageCars = as.factor(train_x$GarageCars)
```

##### Structure of the Data again
```{r}
str(train_x)
```

##### Scale the variables
```{r}
scaling = train_x %>% select(LotFrontage, LotArea, MasVnrArea, BsmtFinSF1, BsmtFinSF2,BsmtUnfSF, TotalBsmtSF, X1stFlrSF, X2ndFlrSF, 
                             LowQualFinSF, GrLivArea, GarageArea, WoodDeckSF, OpenPorchSF, EnclosedPorch, X3SsnPorch, ScreenPorch,
                             PoolArea, MiscVal)

PreNum = preProcess(scaling, method = c('center', 'scale'))
print(PreNum)

train_x = predict(PreNum, train_x)
dim(train_x)

numVars = which(sapply(train_x, is.numeric))
factorVars = which(sapply(train_x, is.factor))

cat('There are', length(numVars), 'numeric variables, and', length(factorVars), 'factor variables')
```
##### Second Correlation Matrix
```{r}
matrix = model.matrix(~0+., data = train_x) %>%
  cor(use = 'pairwise.complete.obs')
matrix_sorted = as.matrix(sort(matrix[, 'SalePrice'], decreasing = T))
MatrixHigh = names(which(apply(matrix_sorted, 1, function(x) abs(x)>0.45)))
cor_factor = matrix[MatrixHigh, MatrixHigh]
corrplot.mixed(cor_factor, tl.col = 'black', tl.pos = 'lt', tl.cex = 0.7, cl.cex = .7, number.cex = .7)
```

##### Drop Variables With Few Distinctions
```{r}
table(train_x$RoofMatl)
```
I decided to drop variables with few distinctions, like RoofMatl. From 2919 observations, 2876 belongs to RoofMatl6, 23 to RoofMatl2, 9 to RoofMatl1 and each of the remaining have only 1. All the variables that have a distribution like this will be dropped from the data.

```{r}
##DROP Street
train_x$Street = NULL

##DROP Condition2
train_x$Condition2 = NULL

##DROP RoofMatl
train_x$RoofMatl = NULL

##DROP Heating
train_x$Heating = NULL

##DROP LowQualFinSF
train_x$LowQualFinSF = NULL

##DROP X3SsnPorch
train_x$X3SsnPorch = NULL

#DROP PoolQC
train_x$PoolQC = NULL

#DROP PoolArea
train_x$PoolArea = NULL
```

### Plots

##### IsNew
```{r}
#IsNew
ggplot(train_x[!is.na(train_x$SalePrice), ], aes(x=IsNew, y=SalePrice)) +
  geom_bar(stat = 'summary', fun ='mean', fill='blue') +
  geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=6) +
  scale_y_continuous(breaks= seq(0, 800000, by=50000)) +
  theme_grey(base_size = 18) +
  geom_hline(yintercept=mean(train$SalePrice), linetype="dashed") + 
  ggtitle('SalePrice by IsNew') + theme(plot.title = element_text(hjust = 0.5))
```

From all the 1460 houses that was sold, 64 of them were new and 1396 were old. The mean value of old houses is almost the same mean from all the houses (dashed line). The mean value from all the houses is almost 190000. New houses have a value around 260000.  

##### BldgType
```{r}
ggplot(train_x[!is.na(train_x$SalePrice), ], aes(x=BldgType, y=SalePrice)) +
  geom_bar(stat = 'summary', fun ='mean', fill='blue') +
  geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=6) +
  scale_y_continuous(breaks= seq(0, 800000, by=50000)) +
  theme_grey(base_size = 18) +
  geom_hline(yintercept=mean(train$SalePrice), linetype="dashed") + 
  ggtitle('SalePrice by BldgType') + theme(plot.title = element_text(hjust = 0.5))
```

##### YearBuilt
```{r}
ggplot(train_x[!is.na(train_x$SalePrice), ], aes(x=YearBuilt, y=SalePrice)) +
  geom_point(color = 'steelblue') + geom_smooth(method = 'lm', color = 'black') +
  scale_y_continuous(breaks= seq(0, 800000, by=50000)) + 
  ggtitle('SalePrice vs YearBuilt Scatterplot') + theme(plot.title = element_text(hjust = 0.5))
```

Like the scatter plot suggests, new houses were sold by higher price then old houses. 

##### TotRoomsAbvGrd
```{r}
ggplot(train_x[!is.na(train_x$SalePrice), ], aes(TotRmsAbvGrd, SalePrice)) + 
  geom_bar(stat = 'summary', fun = 'mean', fill = 'steelblue') +
  geom_label(stat = 'count', aes(label = ..count.., y = ..count..)) +
  scale_y_continuous(breaks = seq(0, 350000, by = 50000)) +
  theme_grey(base_size = 16) +
  geom_hline(yintercept = mean(train$SalePrice), linetype = 'dashed') + 
  ggtitle('Sale Price by Total Rooms Above Ground') + theme(plot.title = element_text(hjust = 0.5))
```

Houses with more rooms are more expensive. While houses with 3 rooms have a mean around 110000, houses with 11 rooms have a mean around 320000. 


##### GrLivArea
```{r}
ggplot(train_x[!is.na(train_x$SalePrice), ], aes(x=GrLivArea, y=SalePrice))+
  geom_point(col='blue') + geom_smooth(method = "lm", se=FALSE, color="black") +
  scale_y_continuous(breaks= seq(0, 800000, by=100000)) +
  geom_text_repel(aes(label = ifelse(GrLivArea > 6, rownames(train_x), ''))) + 
  ggtitle('SalePrice vs YearBuilt Scatterplot') + theme(plot.title = element_text(hjust = 0.5))
```

This plot shows us that as long the ground live area increases, the sale price increases together. Just a reminder, GrLivArea is scaled, thats why the area have this tiny values. Furthermore, the graphic also shows us 2 outliers, even though there are only 2, outliers can have a greatly influence in the model.

##### Check variables in the data set
```{r}
numVars = which(sapply(train_x, is.numeric))
factorVars = which(sapply(train_x, is.factor))

cat('There are', length(numVars), 'numeric variables, and', length(factorVars), 'factor variables')
```

### One Hot Encoding
```{r}
datafor_hot_names = names(factorVars)
data_hot = train_x[, datafor_hot_names]

dummies = dummyVars(" ~. ", data = data_hot)
dummies = data.frame(predict(dummies, newdata = data_hot))
```

After the One Hot Encoding, i choose to drop variables with less than 20 observations. The choice for 20 numbers was because, for example, one variable can have 13 observations in train set and 5 in test set, so i want that every variable have at least 10 observations, so i think 20 is a good number for that condition, although this method is not a savior and can have serious problems. Nothing prevent that a variable with 21 observations can have 19 in the train set and 2 in the test set.  

```{r}
few = which(colSums(dummies[1:nrow(dummies), ], ) <= 20)
fewnames = names(few)
fewnames
```

```{r}
#remove
dummies = dummies[, -few]

numericVarNames = names(numVars)
data_num = train_x[, numericVarNames]

final_data = cbind(dummies, data_num)
```

##### Split the set for Train and Test
```{r}
#Make Train and Test Set again
x.train = data.matrix(final_data[1:1460, 1:252])
x.test = data.matrix(final_data[1461:2919, 1:252])

y.train = final_data$SalePrice[1:1460]
```

##### Ridge Regression
```{r}
ridge.fit = cv.glmnet(x.train, y.train, type.measure = 'mse', alpha = 0, family = 'gaussian')

ridge.pred = predict(ridge.fit, s = ridge.fit$lambda.1se, newx = x.test)
```

##### Lasso Regression
```{r}
lasso.fit = cv.glmnet(x.train, y.train, type.measure = 'mse', alpha = 1, family = 'gaussian')

lasso.pred = predict(lasso.fit, s = lasso.fit$lambda.1se, newx = x.test)
```

##### Elastic Net Regression
```{r}
elastic.fit = cv.glmnet(x.train, y.train, type.measure = 'mse', alpha = 0.5, family = 'gaussian')

elastic.pred = predict(elastic.fit, s = elastic.fit$lambda.1se, newx = x.test)
```

##### Note
That would be the end since i don't have x.test (SalePrice in test set) to make the goodness of fit. I need split to split the train data set, make the regression again and compare the predict with the real values.

```{r, set.seed(4)}
knitr::opts_chunk$set(cache = T)
datasplit = final_data[1:1460, ]

sample = sample.split(datasplit$SalePrice, SplitRatio = 0.7)
train_for = subset(datasplit, sample == T)
test_for = subset(datasplit, sample == F)

x_train = data.matrix(train_for[, 1:252])
x_test = data.matrix(test_for[, 1:252])

y_train = train_for$SalePrice
y_test = test_for$SalePrice
```

##### Ridge
```{r}
alpha0.fit = cv.glmnet(x_train, y_train, type.measure = 'mse', alpha = 0, family = 'gaussian')

pred.alpha0 = predict(alpha0.fit, s = alpha0.fit$lambda.min, newx = x_test)
```

##### SSE and R^2^
```{r}
sst = sum((y_test - mean(y_test))^2)
sse.alpha0 = sum((pred.alpha0 - y_test)^2)
mean((y_test - pred.alpha0)^2)

#find R-Squared
rsq.alpha0 <- 1 - sse.alpha0/sst
rsq.alpha0
```
Ridge Regression model had a R^2^ of 0.8763.  

##### Lasso
```{r}
alpha1.fit = cv.glmnet(x_train, y_train, type.measure = 'mse', alpha = 1, family = 'gaussian')

pred.alpha1 = predict(alpha1.fit, s = alpha1.fit$lambda.min, newx = x_test)
```

##### SSE and R^2^
```{r}
sse.alpha1 = sum((pred.alpha1 - y_test)^2)
mean((y_test - pred.alpha1)^2)

#find R-Squared
rsq.alpha1 <- 1 - sse.alpha1/sst
rsq.alpha1
```
Lasso Regression model had a R^2^ of 0.8774.

##### Elastic Net
```{r}
alpha0.5.fit = cv.glmnet(x_train, y_train, type.measure = 'mse', alpha = 0.5, family = 'gaussian')

pred.alpha0.5 = predict(alpha0.5.fit, s = alpha0.5.fit$lambda.min, newx = x_test)
```

##### SSE and R^2^
```{r}
sse.alpha0.5 = sum((pred.alpha0.5 - y_test)^2)
mean((y_test - pred.alpha0.5)^2)

#find R-Squared
rsq.alpha0.5 <- 1 - sse.alpha0.5/sst
rsq.alpha0.5
```
Elastic Net Regression had a R^2^ of 0.8770.

Ridge, Lasso and Elastic Net regressions get almost the same results. While I was testing the model and before the seed was settled, these three regressions were changing positions of the best model possible.  

On the chunk below i will find the best alpha that i can put in the model.

##### Finding the best alpha
```{r}
list.of.fits = list()
for (i in 0:10) {
  model.name = paste0("alpha", i/10)
  
  list.of.fits[[model.name]] = 
    cv.glmnet(x_train, y_train, type.measure = 'mse', alpha = i/10, family = 'gaussian')
}

results = data.frame()
for (i in 0:10) {
  model.name = paste0("alpha", i/10)
  
  predicted = predict(list.of.fits[[model.name]], s = list.of.fits[[model.name]]$lambda.min, newx = x_test)
  mse = mean((y_test - predicted)^2)
  temp = data.frame(alpha = i/10, mse = mse, model.name = model.name)
  results = rbind(results, temp)
}

results
```

We can see that again, the MSE of the models were too close between them, but the best alpha was 0.1. It's important to know that the results can change drastically depending on sample to sample. Some samples is easier to predict and get less MSE and some samples are harder to predict and your MSE will be high.  


##### Linear Model
```{r}
linear.fit = lm(SalePrice ~., data = train_for)

pred.linear = predict(linear.fit, newdata = test_for)

sse.linear = sum((test_for$SalePrice - pred.linear)^2)
mean((test_for$SalePrice - pred.linear)^2)
sst.linear = sum((test_for$SalePrice - mean(test_for$SalePrice))^2)

rsq.linear = 1 - sse.linear/sst.linear
rsq.linear
```

Linear Regression had a R^2^ of 0.7639.





*Dataset:*<https://www.kaggle.com/competitions/house-prices-advanced-regression-techniques/data>  
